```{r, echo = FALSE, message = FALSE}
source(here::here("scripts/setup.R"))
```

# 4. <strong>EDA</strong>

## 4.1 Evolution of shark attacks through the years

With the y-axis showing the frequency of shark attacks and the x-axis showing the passage of time, the graph below displays a striking pattern in shark attacks over time. The data exhibits a recognizable pattern that indicates a steady increase in the quantity of shark attacks within the designated period. This rising tendency begs critical concerns regarding what is causing such an increase and necessitates more research into ecological, environmental, and human-related elements that could be involved in this occurrence. The graph emphasizes the need of keeping an eye on and comprehending the dynamics of these episodes in addition to providing a visual depiction of the rise in shark attacks. 

This trend is a key research topic: why are marine attacks increasing year after year? What could be the reasons? In fact, the reasons can be various, for example of an ecological, environmental nature or linked to human seaside activity. The graph highlights the need to keep an eye on and understand the dynamics of these episodes, which is exactly what we will do during our study.

Three fundamental areas of this graph should be an element of analysis: 1975-1978; 1988-1990; 2019-2020. In fact, we could ask ourselves, why is there a decrease in attacks in this period?
For the last period, the answer is simple: our dataset presents data up to June 2018, for the second half of the year we have no data, which explains the drastic drop in attacks.
For the other two periods, we will discover the reasons during a more in-depth analysis

```{r, echo=TRUE}
plot1 <- ggplot(data = attacks, aes(x = Year)) +
  geom_smooth(stat = "count", aes(group = 1), color = "blue", size = 1) +
  geom_bar(stat = "count", fill = "red") +
  labs(title = "Shark Attacks Evolution Over Years", x = "Year", y = "Number of Attacks")

plot1
```

## 4.2 Pattern of shark attacks frequencies

<strong> Pattern of shark attacks frequencies through the day</strong> 

The bar graph below, where the x-axis indicates four time categories —morning, afternoon, evening, and night— displays the distribution of shark attacks throughout the day. The graph's most noticeable aspect is how frequently shark attacks occur in the afternoon, followed by the morning, evening and night. This concentration of incidents in the afternoon raises intriguing questions about potential human behavioral factors that could contribute to this temporal pattern. 

First and foremost, we believe that the afternoon is when most people choose to go swimming, which contributes to the frequency of shark attacks during this time. Recreational water activities such as swimming, surfing, kayaking, and other water sports are probably more popular during these times. Furthermore, the afternoon is when the seas are the warmest, so it's intriguing to find out if this aspect also has a significant influence on shark activity.
Conversely, the decreased number of assaults that occur at night may be related to less people using the water during that time, but there may be other possible reasons as well (less high temperatures, poorer visibility).

Lastly, we suppose that the medium frequencies of the morning and evening might correspond to transitional periods when shark and human behavior are changing.

```{r, echo=TRUE}

bar_colors <- c("#66c2a5", "#fc8d62", "#8da0cb", "#e78ac3")

# Create a sorted table
sorted_table <- table(attacks$Time)
sorted_table <- sorted_table[order(-sorted_table)]

# Create a bar plot with the sorted data
barplot1 <- barplot(sorted_table,
                    col = bar_colors,
                    xlab = "Time of Day", ylab = "Frequency of Attacks",
                    main = "Frequency of Shark Attacks by Time of Day",
                    border = "white",
                    ylim = c(0, 1800),
                    space = 0.5,
                    cex.names = 0.8,
                    font.axis = 2,
                    beside = TRUE)


# Add text labels with frequencies on the bars
text_labels <- sorted_table
text(x = barplot1, y = sorted_table, labels = text_labels, pos = 3, cex = 0.8, col = "black")
```

<strong>Pattern of shark attacks frequencies per country</strong> 

A first glance at our plotted data reveals that the United States, Australia, and South Africa are the three primary countries with a significant number of shark attacks. 

```{r, echo=TRUE}
shark_attacks_by_country <- attacks %>%
  group_by(Country) %>%
  summarise(total_attacks = n())

shark_attacks_by_country <- shark_attacks_by_country %>%
  mutate(CountryCategory = ifelse(total_attacks >= 30, as.character(Country), "Other"))

# Order the countries by frequency in descending order
shark_attacks_by_country$CountryCategory <- reorder(shark_attacks_by_country$CountryCategory, -shark_attacks_by_country$total_attacks)

# Plot the data
plot2<- ggplot(data = shark_attacks_by_country, aes(x = total_attacks, y = CountryCategory)) +
  geom_col(fill = "skyblue") +
  labs(title = "Total Shark Attacks in Each Country", x = "Number of Attacks", y = "Country") +
  theme_minimal() +
  theme(axis.text.y = element_text(hjust = 1)) +
  scale_y_discrete(labels = scales::label_wrap(10))

interactive_plot <- ggplotly(plot2)
interactive_plot

```

## 4.3 Fatality of shark attacks

<strong>Are shark attacks more likely to lead to death?</strong> 

IT WAS A PIECHART...

<strong>Fatalities by type of shark</strong> 

As we can see in this graph, most of attacks were not fatal, meaning that sharks injured a lot of people without taking their lives away. The types of sharks that did the most of accidents were the following ones: Bull shark, Tiger shark and White shark, with the deadlies being the last one. Why do they commit attacks? Probably because their feeding behavior. Maybe they misidentify their predators attacking people instead of animals like turtles, seals, or sea lions. This is consistent with Bull shark and Tiger shark being repetitevly named as the deadlies sharks in the seas and oeceans (CBS News, 2010).
Another relevant information derived from this graph is the big number of unidentified sharks. Out of all the attacks recorded by our dataset, we have almost 600 Unidentified species. It is such a pity not to be able to have this information, which would have led us to more insightful results.


```{r, echo=TRUE}
summary_table <- table(attacks$Species, attacks$Fatality)

summary_df <- as.data.frame(summary_table)
plot4 <- ggplot(summary_df, aes(x = Var1, y = Freq, fill = Var2)) +
  geom_bar(stat = "identity") +
  labs(
    title = "Fatality by Type of Shark",
    x = "Type of Shark",
    y = "Frequency",
    fill = "Legend"
  ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_fill_manual(
    name = "Was the shark attack fatal?",
    values = c("blue", "pink", "yellow"),
    labels = c("Yes", "No", "Unknown")
  )

print(plot4)
```

## 4.4 Evolution of temperatures through the years

The graph below illustrates the change in global temperature throughout the last few decades, showing an overall increase in temperature over the 50-year period. Despite some fluctuations, the upward trend indicates a clear warming pattern, with a notable acceleration following the 1980s and one of the two most significant rises occurring in the recent decade.


```{r, echo=TRUE}
# Filter out NA values and non-numeric values in the Temperature column
temperature_data_filtered <- temperature %>%
  filter(!is.na(Temperature), !is.na(as.numeric(Temperature)))

# Convert the Temperature column to numeric
temperature_data_filtered$Temperature <- as.numeric(temperature_data_filtered$Temperature)

# Calculate the mean temperature for each year
world_temperature <- temperature_data_filtered %>%
  group_by(Year) %>%
  summarise(World_Temperature = mean(Temperature, na.rm = TRUE))

# Create a line plot for the world temperature
plot5 <- ggplot(world_temperature, aes(x = Year, y = World_Temperature)) +
  geom_line() +
  labs(title = "World Temperature Change Over Years",
       x = "Year",
       y = "World Temperature Change") +
  theme_minimal()

interactive_plot <- ggplotly(plot5)
interactive_plot

```


## 4.5 Evolution of sea level through the yars

This line graph depicts the trend of rising sea levels from around the ‘90s until the early 2020s, showing some fluctuations initially, and a notable upward trend thereafter. The early 2000s represent a steady rate of increase, with a minor decrease in the early 2010s. This pattern appears to be somewhat consistent with global temperature changes, and a steepening rise in the last couple of years indicate a complementary relation with the increasing temperature levels.

```{r, echo=TRUE}
# Create a line plot
plot6 <- ggplot(data = sealevel, aes(x = Year, y = GMSL_GIA)) +
  geom_line(color = "green", size = 1.5) +
  labs(title = "Evolution of Sea Level Over the Years", x = "Year", y = "Sea Level")

interactive_plot <- ggplotly(plot6)
interactive_plot
```

## 4.6 CO2 emissions insights

<strong>Trend of CO2 emissions per year</strong> 

The bar graph reveals an evident increase of emissions in the depicted period, as indicated by a consistent upward trend with few fluctuations and no significant declines. The rise reflects the growth of industrialisation over the years, especially marked by the end of the Cold War and subsequent economic growth in many parts of the world. The early 2000s show a particularly sharp increase, which is likely to be as a result of the rapid increase in manufacturing and energy production in Asian regions. Post-2010 years show some fluctuations that could be attributed to emissions mitigation efforts, but the trend overall remains the same.


```{r, echo=TRUE}
# Create a line plot with backticks to show the trend in co2 emissions throughout the years
plot7 <- ggplot(data = co2, aes(x = Year, y = `Annual CO2 Emissions`)) +
  geom_line(color = "blue") +
  labs(title = "Trend of CO2 Emissions Over the Years", x = "Year", y = "Annual CO2 Emissions")

interactive_plot <- ggplotly(plot7)
interactive_plot
```

<strong>Which continent generates more GHG emissions? </strong> 

We can see that there is a significant disparity in emissions among different continents; Asia shows a sharp increase in emissions over the years, particularly accelerating after 2000, becoming the continent with the highest CO2 emissions. This is perfectly consistent with other data, showing that 5 out of the 6 most polluting countries in the world are in Asian, namely China, India, Japan, Russia and Iran. (Statista, 2021)
Emissions in Europe depict a gradual rise until the early 90’s, with a noticeable fluctuation and decline thereafter. 
The North American region follows a similar trend to Europe with a steady increase until the 2000s, followed by a levelling off and a decline. 
On the other hand, CO2 emissions from South America show a steady increase over time with less fluctuations as compared to Europe and NA.
Africa’s emissions have additionally been increasing at a steady rate over the last few decades, but they remain significantly lower compared to the rest of the continents. 
Oceania represents the lowest overall emissions among the listed regions, with an increase rate on par with Africa.
Antarctica’s emissions are not visible on the graph, which is consistent with expectations as it is a largely uninhabited continent.


```{r, echo=TRUE, results='hide', warning=FALSE, message=FALSE}

co2_grouped <- co2 %>%
  mutate(Region = case_when(
    ISO_Code %in% c(NA) ~ "Unknown",
    ISO_Code %in% c("AFG", "ARM", "AZE", "BHR", "BGD", "BTN", "BOL", "BGR", "BFA", "BDI", "KHM", "CMR", "CPV", "CHN", "CXR", "CYP", "HKG", "IND", "IDN", "IRN", "IRQ", "ISR", "JOR", "JPN", "KAZ", "KWT", "KGZ", "LAO", "LBN", "MAC", "MYS", "MDV", "MNG", "MMR", "NPL", "PRK", "OMN", "PAK", "PSE", "PHL", "QAT", "SAU", "SGP", "KOR", "LKA", "TUR", "SYR", "TWN", "TJK", "THA", "TKM", "ARE", "UZB", "YEM", "VNM") ~ "Asia",
    ISO_Code %in% c("ALB", "AND", "AUT", "BLR", "BEL", "BIH", "BWA", "HRV", "CZE", "DNK", "EST", "FIN", "FRO", "FRA", "GEO", "DEU", "GRC", "GRL", "GLP", "HUN", "ISL", "IRL", "ITA", "LVA", "LTU", "LIE", "LUX", "MLT", "MDA", "MNE", "NLD", "MKD", "NOR", "POL", "PRT", "ROU", "RUS", "SRB", "SVK", "SVN", "ESP", "SWE", "CHE", "UKR", "GBR") ~ "Europe",
    ISO_Code %in% c("DZA", "AGO", "BEN", "CAF", "TCD", "COM", "COG", "CIV", "COD", "DJI", "TLS", "EGY", "ERI", "SWZ", "ETH", "GNQ", "GAB", "GMB", "GHA", "GIN", "GNB", "KEN", "LSO", "LBR", "LBY", "MDG", "MWI", "MLI", "MTQ", "MRT", "MUS", "MYT", "MAR", "MOZ", "NAM", "NER", "NGA", "REU", "RWA", "SHN", "SEN", "SYC", "SLE", "STP", "ZAF", "SOM", "SSD", "SDN", "SUR", "TZA", "TGO", "TUN", "UGA", "ZMB", "ZWE") ~ "Africa",
    ISO_Code %in% c("AIA", "ATG", "ABW", "BHS", "BMU", "BRB", "BLZ", "VGB", "BRN", "CAN", "CRI", "CUB", "CUW", "DMA", "DOM", "SLV", "GRD", "GTM", "HTI", "HND", "JAM", "MSR", "MEX", "USA", "NIC", "PAN", "KNA", "LCA", "SPM", "VCT", "SXM", "TCA", "TTO") ~ "North America",
    ISO_Code %in% c("ARG", "COL", "CHL", "ECU", "GUF", "GUY", "PRY", "PER", "URY", "VEN", "BRA", "BES") ~ "South America",
    ISO_Code %in% c("AUS", "COK", "FJI", "PYF", "KIR", "MHL", "FSM", "NRU", "NCL", "NZL", "NIU", "PLW", "PNG", "WSM", "SLB", "TON", "TUV", "VUT", "WLF") ~ "Australia & Oceania",
    ISO_Code %in% c("ATA") ~ "Antarctica"
  ))


# Step 1: Calculate Mean Emissions by Year and Region
mean_emissions <- co2_grouped %>%
  group_by(Year, Region) %>%
  summarize(Mean_Emissions = mean(`Annual CO2 Emissions`, na.rm = TRUE))


# Now we get rid of that Unknown Region to have a better view of the other Regions

# Step 1: Calculate Mean Emissions by Year and Region and getting rid of the things we don't want
mean_emissions <- co2_grouped %>%
  filter(!is.na(Region) & !is.na(`ISO_Code`) & Region != "Unknown") %>%
  group_by(Year, Region) %>%
  summarize(Mean_Emissions = mean(`Annual CO2 Emissions`, na.rm = TRUE))


```

```{r, echo=TRUE}
# Step 2: Let's plot the Data
plot2 <- ggplot(mean_emissions, aes(x = Year, y = Mean_Emissions, color = Region)) +
  geom_line() +
  labs(title = "CO2 Emissions by Continent",
       x = "Year",
       y = "Annual CO2 Emissions") +
  theme_minimal()

interactive_plot <- ggplotly(plot2)
interactive_plot
```


Shows the relationship between the attacks and the age of the victims 

```{r, echo=TRUE}

# Create a summary table with the count of attacks for each age
attacks_summary <- attacks %>%
  group_by(Age) %>%
  summarise(NumAttacks = n())

# Scatter plot to show the relationship between the number of shark attacks and age
interactive_attacks_age <- ggplotly(
  ggplot(attacks_summary, aes(x = Age, y = NumAttacks)) +
    geom_point(color = "blue", size = 2) +
    geom_smooth(method = "lm", se = FALSE, color = "red") +  # Add a linear trend line
    labs(title = "Relationship between Shark Attacks and Age",
         x = "Age",
         y = "Number of Attacks")
)

# Display the interactive scatter plot
print(interactive_attacks_age)

```

Better understanding of the attacks and sex of the victims


```{r, echo=TRUE}
# Create a summary table with the count of attacks for each sex
attacks_summary_sex <- attacks %>%
  group_by(Sex) %>%
  summarise(NumAttacks = n())

# Define the desired order of levels
sex_order <- c("M", "F", "Unknown")

# Convert the 'Sex' variable to a factor with the specified order
attacks_summary_sex$Sex <- factor(attacks_summary_sex$Sex, levels = sex_order)

# Bar plot to show the distribution of shark attacks by sex
interactive_bar_plot_sex_attacks <- ggplotly(
  ggplot(attacks_summary_sex, aes(x = Sex, y = NumAttacks, fill = Sex)) +
    geom_bar(stat = "identity", position = "dodge", width = 0.7) +
    labs(title = "Distribution of Shark Attacks by Sex",
         x = "Sex",
         y = "Number of Attacks") +
    scale_fill_manual(values = c("blue", "pink", "gray"))
)

# Display the interactive bar plot
print(interactive_bar_plot_sex_attacks)

```

Want to know where are the regions (continents) with the most attacks

```{r, echo=TRUE}

#Create a new column in attacks that represents the continents
attacks_grouped <- attacks %>%
  mutate(Region = case_when(
    ISO_Code %in% c(NA) ~ "Unknown",
    ISO_Code %in% c("AFG", "ARM", "AZE", "BHR", "BGD", "BTN", "BOL", "BGR", "BFA", "BDI", "KHM", "CMR", "CPV", "CHN", "CXR", "CYP", "HKG", "IND", "IDN", "IRN", "IRQ", "ISR", "JOR", "JPN", "KAZ", "KWT", "KGZ", "LAO", "LBN", "MAC", "MYS", "MDV", "MNG", "MMR", "NPL", "PRK", "OMN", "PAK", "PSE", "PHL", "QAT", "SAU", "SGP", "KOR", "LKA", "TUR", "SYR", "TWN", "TJK", "THA", "TKM", "ARE", "UZB", "YEM", "VNM") ~ "Asia",
    ISO_Code %in% c("ALB", "AND", "AUT", "BLR", "BEL", "BIH", "BWA", "HRV", "CZE", "DNK", "EST", "FIN", "FRO", "FRA", "GEO", "DEU", "GRC", "GRL", "GLP", "HUN", "ISL", "IRL", "ITA", "LVA", "LTU", "LIE", "LUX", "MLT", "MDA", "MNE", "NLD", "MKD", "NOR", "POL", "PRT", "ROU", "RUS", "SRB", "SVK", "SVN", "ESP", "SWE", "CHE", "UKR", "GBR") ~ "Europe",
    ISO_Code %in% c("DZA", "AGO", "BEN", "CAF", "TCD", "COM", "COG", "CIV", "COD", "DJI", "TLS", "EGY", "ERI", "SWZ", "ETH", "GNQ", "GAB", "GMB", "GHA", "GIN", "GNB", "KEN", "LSO", "LBR", "LBY", "MDG", "MWI", "MLI", "MTQ", "MRT", "MUS", "MYT", "MAR", "MOZ", "NAM", "NER", "NGA", "REU", "RWA", "SHN", "SEN", "SYC", "SLE", "STP", "ZAF", "SOM", "SSD", "SDN", "SUR", "TZA", "TGO", "TUN", "UGA", "ZMB", "ZWE") ~ "Africa",
    ISO_Code %in% c("AIA", "ATG", "ABW", "BHS", "BMU", "BRB", "BLZ", "VGB", "BRN", "CAN", "CRI", "CUB", "CUW", "DMA", "DOM", "SLV", "GRD", "GTM", "HTI", "HND", "JAM", "MSR", "MEX", "USA", "NIC", "PAN", "KNA", "LCA", "SPM", "VCT", "SXM", "TCA", "TTO") ~ "North America",
    ISO_Code %in% c("ARG", "COL", "CHL", "ECU", "GUF", "GUY", "PRY", "PER", "URY", "VEN", "BRA", "BES") ~ "South America",
    ISO_Code %in% c("AUS", "COK", "FJI", "PYF", "KIR", "MHL", "FSM", "NRU", "NCL", "NZL", "NIU", "PLW", "PNG", "WSM", "SLB", "TON", "TUV", "VUT", "WLF") ~ "Australia & Oceania",
    ISO_Code %in% c("ATA") ~ "Antarctica"
  ))


# Filter data to include only valid regions (excluding "Unknown")
filtered_attacks_region <- attacks_grouped %>% 
  filter(!is.na(Region) & Region != "Unknown")

# Bar plot to show the distribution of shark attacks by region
interactive_bar_plot_region_attacks <- ggplotly(
  ggplot(filtered_attacks_region, aes(x = Region, fill = Region)) +
    geom_bar() +
    labs(title = "Distribution of Shark Attacks by Region",
         x = "Region",
         y = "Number of Attacks") +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) # Rotate x-axis labels for better readability
)


# Display the interactive bar plot
print(interactive_bar_plot_region_attacks)  


```

Does sea level and the number of shark attacks have a relationship?


```{r, echo=TRUE}
# Calculate the number of attacks per year
attacks_summary <- attacks %>%
  group_by(Year) %>%
  summarize(Number_of_Attacks = n())

# Assuming there's a common column named "Year" in both datasets
merged_data5 <- merge(attacks_summary, sealevel, by = "Year", all.x = TRUE)

correlation_coefficient <- cor(merged_data5$Number_of_Attacks, merged_data5$GMSL_GIA, use = "complete.obs")

# Scatter plot
interactive_scatter_plot_sea_level_attacks <- ggplotly(
  ggplot(merged_data5, aes(x = GMSL_GIA, y = Number_of_Attacks)) +
    geom_point() +
    geom_smooth(method = "lm", se = FALSE, color = "red") +  # Add a linear trend line
    labs(title = "Relationship Between Sea Level and Shark Attacks",
         x = "Global Mean Sea Level Global Isostatic Adjustment (GMSL_GIA)",
         y = "Number of Attacks")
)

# Display the interactive scatter plot
print(interactive_scatter_plot_sea_level_attacks)


```

Relationship between temperature change and the number of shark attacks

```{r, echo=TRUE}


# Identify non-numeric values in the Temperature column
non_numeric_temp <- temperature %>%
  filter(!is.numeric(Temperature)) %>%
  distinct(Temperature)

# Convert "Temperature" column to numeric
temperature$Temperature <- as.numeric(temperature$Temperature)

# Check for negative values
negative_values <- temperature %>%
  filter(Temperature < 0)

# Calculate the mean temperature
mean_temp_world <- temperature %>%
  group_by(Year) %>%
  summarize(Mean_Temperature = mean(Temperature, na.rm = TRUE))

# Filter out non-numeric values in the Temperature column
temperature2 <- temperature %>%
  filter(is.numeric(Temperature))

# Calculate mean temperature for the world per year
mean_temp_world <- temperature2 %>%
  group_by(Year) %>%
  summarize(Mean_Temperature = mean(Temperature, na.rm = TRUE))

# Merge datasets
merged_data6 <- merge(attacks_summary, mean_temp_world, by = "Year", all.x = TRUE)

# Plotting
interactive_scatter_plot_worldtemperature_attacks <- ggplotly(
  ggplot(merged_data6, aes(x = Mean_Temperature, y = Number_of_Attacks)) +
    geom_point() +
    geom_smooth(method = "lm", se = FALSE, color = "red") +  # Add a linear trend line
    labs(title = "Relationship between Shark Attacks and World Mean Temperature Change",
         x = "World Mean Temperature Change",
         y = "Number of Shark Attacks") +
    theme_minimal()
)


print(interactive_scatter_plot_worldtemperature_attacks)

```

What are the activities with the most attacks

```{r, echo=TRUE}

# Check the top activities with the most attacks
top_activities <- attacks %>%
  group_by(Activity, Sex) %>%
  summarize(Number_of_Attacks = n()) %>%
  arrange(desc(Number_of_Attacks)) %>%
  top_n(30)  # Adjust the number if you want more or fewer top activities

# Plot the results
top_activities_attacks <- ggplot(top_activities, aes(x = reorder(Activity, -Number_of_Attacks), y = Number_of_Attacks, fill = Sex)) +
  geom_bar(stat = "identity") +
  labs(title = "Top Activities with the Most Attacks",
       x = "Activities",
       y = "Number of Attacks",
       fill = "Sex") +
  scale_fill_manual(values = c("pink", "blue", "orange"), name = "Sex") +
  theme_minimal() + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))


interactive_top_activities_attacks <- ggplotly(top_activities_attacks)
interactive_top_activities_attacks


```

See the evolution of shark attacks throughout the years with sex DONE


```{r, echo=TRUE}

# Create a plot to show the trend of shark attacks throughout the years, including victim's sex (interactive version)
attacks_evolution_sex <- ggplot(data = attacks, aes(x = Year, fill = Sex)) +
  geom_bar(position = "stack", color = "white") +
  labs(title = "Shark Attacks Evolution Over Years by Sex",
       x = "Year",
       y = "Number of Attacks",
       fill = "Sex") +
  scale_fill_manual(values = c("pink", "blue", "orange"), name = "Sex") +
  theme_minimal()


interactive_attacks_evolution_sex <- ggplotly(attacks_evolution_sex)
interactive_attacks_evolution_sex

```

Shark attacks monthly repartition



```{r, echo=TRUE}


#Here it is great but I tried to class the months in order but it has not worked yet

# Create a plot to show the repartition of shark attacks per month by sex
attacks_per_months <- ggplot(data = attacks, aes(x = factor(Date), fill = Sex)) +
  geom_bar(position = "stack", color = "white") +
  labs(title = "Shark Attacks Repartition by Month and Sex",
       x = "Month",
       y = "Number of Attacks",
       fill = "Sex") +
  scale_fill_manual(values = c("pink", "blue", "orange"), name = "Sex") +
  theme_minimal()


interactive_attacks_per_months <- ggplotly(attacks_per_months)
interactive_attacks_per_months

```


```{r, echo=TRUE}



```



```{r, echo=TRUE}



```



```{r, echo=TRUE}



```
